---
title: "Code Report"
date: "2018/5/9"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Library
```{r cars}
rm(list=ls())
options(warn=-1)
## Library ##
library(randomForest)
library(randomForestExplainer)
library(corrplot)
library(rpart)
library(knitr)
library(MASS)
library(factoextra)
library(dplyr)
library(kableExtra)
library(caret)
library(glmnet)
library(ggplot2)
library(plotmo)
library(cvTools)
library(data.table)
```

## Function for cross-validation

* ### Model
    + randomforest
    + ridge
    + lasso
* ### methods
    + directly
    + two-step


```{r}
## Functions ##
crossvalidate<-function(data, Model, methods){
  # data is the training set with the "seasons" column
  # Model is a string describing a model
  # methods is a string with the name of predicting type we want

  # Initialize empty list for recording performances
  performances<-c()
  k<-length(unique(data$seasons))
  # One iteration per fold
  for (fold in 1:k)
  {
    # Create training set for this iteration
    # Subset all the datapoints where seasons does not match the current fold
    training_set <- data[data$seasons != fold,]
    # Create test set for this iteration
    # Subset all the datapoints where seasons matches the current fold
    testing_set <- data[data$seasons == fold,]
    ## Train model
    if(Model=="randomforest")
    {
      if(methods=="directly")
      {
        model<-randomForest(training_set[,-(1:4)],training_set$rank_in_finals,ntree=500)
        predicted<-predict(model,newdata=testing_set[,-(1:4)])
        predicted[rank(predicted)>3]<-10
        MSPE<-mspe(testing_set$rank_in_finals,predicted)
        #RMSPE<-rmspe(testing_set$rank_in_finals,predicted)
        #MAPE<-mape(testing_set$rank_in_finals,predicted)
        rp<-rank(predicted,ties.method="first")
        rts<-rank(testing_set$rank_in_finals,ties.method="first")
        corders<-sum(rts[which(rp<=3)]==rp[which(rp<=3)])
        performances<-rbind(performances,c(MSPE=MSPE,correct_count=corders))
      }
      if(methods=="two-step")
      {
        model<-randomForest(training_set[,-(1:4)],training_set$top3_or_not,ntree=500)
        trained<-training_set$top3_or_not
        predicted<-predict(model,newdata=testing_set[,-(1:4)])
        predicted<-as.numeric(rank(-predicted)<=3)
        model<-randomForest(cbind(training_set[,-(1:4)],new=trained),
                            training_set$rank_in_finals,ntree=500)
        predicted<-predict(model,newdata=cbind(testing_set[,-(1:4)],new=predicted))
        predicted[rank(predicted)>3]<-10
        MSPE<-mspe(testing_set$rank_in_finals,predicted)
        #RMSPE<-rmspe(testing_set$rank_in_finals,predicted)
        #MAPE<-mape(testing_set$rank_in_finals,predicted)
        rp<-rank(predicted,ties.method="first")
        rts<-rank(testing_set$rank_in_finals,ties.method="first")
        corders<-sum(rts[which(rp<=3)]==rp[which(rp<=3)])
        performances<-rbind(performances,c(MSPE=MSPE,correct_count=corders))
      }
    }
    if(Model=="ridge")
    {
      if(methods=="directly")
      {
        model<-cv.glmnet(x=as.matrix(training_set[,-(1:4)]),y=training_set$rank_in_finals,
                         type.measure="mse",nfolds=4,alpha=0)
        predicted<-predict(model,new=as.matrix(testing_set[,c(-(1:4))]))
        predicted[rank(predicted)>3]<-10
        MSPE<-mspe(testing_set$rank_in_finals,predicted)
        #RMSPE<-rmspe(testing_set$rank_in_finals,predicted)
        #MAPE<-mape(testing_set$rank_in_finals,predicted)
        rp<-rank(predicted,ties.method="first")
        rts<-rank(testing_set$rank_in_finals,ties.method="first")
        corders<-sum(rts[which(rp<=3)]==rp[which(rp<=3)])
        performances<-rbind(performances,c(MSPE=MSPE,correct_count=corders))
      }
      if(methods=="two-step")
      {
        model<-cv.glmnet(x=as.matrix(training_set[,-(1:4)]),y=training_set$top3_or_not,
                         type.measure="mse",nfolds=4,alpha=0)
        trained<-training_set$top3_or_not
        predicted<-predict(model,new=as.matrix(testing_set[,-(1:4)]))
        predicted<-as.numeric(rank(-predicted)<=3)
        model<-cv.glmnet(x=as.matrix(cbind(training_set[,-(1:4)],new=trained)),y=training_set$rank_in_finals,
                         type.measure="mse",nfolds=4,alpha=0)
        predicted<-predict(model,new=as.matrix(cbind(testing_set[,-(1:4)],new=predicted)))
        predicted[rank(predicted)>3]<-10
        MSPE<-mspe(testing_set$rank_in_finals,predicted)
        #RMSPE<-rmspe(testing_set$rank_in_finals,predicted)
        #MAPE<-mape(testing_set$rank_in_finals,predicted)
        rp<-rank(predicted,ties.method="first")
        rts<-rank(testing_set$rank_in_finals,ties.method="first")
        corders<-sum(rts[which(rp<=3)]==rp[which(rp<=3)])
        performances<-rbind(performances,c(MSPE=MSPE,correct_count=corders))
      }
    }
    if(Model=="lasso")
    {
      if(methods=="directly")
      {
        model<-cv.glmnet(x=as.matrix(training_set[,-(1:4)]),y=training_set$rank_in_finals,
                         type.measure="mse",nfolds=4,alpha=1)
        predicted<-predict(model,new=as.matrix(testing_set[,c(-(1:4))]))
        predicted[rank(predicted)>3]<-10
        MSPE<-mspe(testing_set$rank_in_finals,predicted)
        #RMSPE<-rmspe(testing_set$rank_in_finals,predicted)
        #MAPE<-mape(testing_set$rank_in_finals,predicted)
        rp<-rank(predicted,ties.method="first")
        rts<-rank(testing_set$rank_in_finals,ties.method="first")
        corders<-sum(rts[which(rp<=3)]==rp[which(rp<=3)])
        performances<-rbind(performances,c(MSPE=MSPE,correct_count=corders))
      }
      if(methods=="two-step")
      {
        model<-cv.glmnet(x=as.matrix(training_set[,-(1:4)]),y=training_set$top3_or_not,
                         type.measure="mse",nfolds=4,alpha=1)
        trained<-training_set$top3_or_not
        predicted<-predict(model,new=as.matrix(testing_set[,-(1:4)]))
        predicted<-as.numeric(rank(-predicted)<=3)
        model<-cv.glmnet(x=as.matrix(cbind(training_set[,-(1:4)],new=trained)),y=training_set$rank_in_finals,
                         type.measure="mse",nfolds=4,alpha=1)
        predicted<-predict(model,new=as.matrix(cbind(testing_set[,-(1:4)],new=predicted)))
        predicted[rank(predicted)>3]<-10
        MSPE<-mspe(testing_set$rank_in_finals,predicted)
        #RMSPE<-rmspe(testing_set$rank_in_finals,predicted)
        #MAPE<-mape(testing_set$rank_in_finals,predicted)
        rp<-rank(predicted,ties.method="first")
        rts<-rank(testing_set$rank_in_finals,ties.method="first")
        corders<-sum(rts[which(rp<=3)]==rp[which(rp<=3)])
        performances<-rbind(performances,c(MSPE=MSPE,correct_count=corders))
      }
    }
  }
  return(performances)
}

plotCoeffEvolution=function(penalizedGlm)
{
  lambda=penalizedGlm$lambda
  coeff=as.matrix(penalizedGlm$beta)
  rowName=rownames(coeff)
  coeff=data.table(coeff)
  coeff[,name:=rowName]
  coeff=melt(coeff,id.vars = 'name')
  coeff[,variable:=rep(lambda,each=length(unique(name)))]
  ggplot(coeff,aes(x=variable,y=value,color=name))+
    geom_line()+
    xlab('lambda')+
    ylab('Value of coefficient')+
    scale_x_log10()+
    theme_minimal()
}
```

## Data Cleaning

```{r eval=T}
## Cleaning ##
data<-read.table("data.csv",sep=",",header=T,stringsAsFactors=F)
colnames(data)<-c("name","rank_in_finals","top3_or_not","final_or_not","seasons",
                  "sex","no_of_top3","no_of_top1","in_group","start","time",
                  "professional","mean","median","mode","std","regular","best","worst")
data[is.na(data)]<-0
data$sex<-ifelse(data$sex=="男",1,0);data$rank_in_finals<-ifelse(!data$rank_in_finals,10,data$rank_in_finals)
data$final_or_not<-NULL
```

## Correlation Matrix

```{r eval=T}
## Correlation Matrix ##
corrplot(cor(data[,-(1:4)]),order="hclust",addrect = 12)
```

#### Mean, median and mode are highly correlated.
#### Drop median as well as mode before building our models.

```{r eval=T}
## Drop mode and median ##
data$median<-data$mode<-NULL
```

```{r eval=T}
## PCA ##
pca<-prcomp(~.,data[,-(1:4)])
summary(pca)
fviz_eig(pca,addlabels = T)
fviz_pca_var(pca, col.var = "contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), 
             repel = TRUE) # Avoid text overlapping
```

## Cross Validation Results
```{r eval=T}
data_train<-data[data$seasons<=5,];data_test<-data[data$seasons==6,]
k<-crossvalidate(data_train,Model="randomforest",methods="directly")
cat("Model: randomforest\t\tMethods: directly\n\n");k;cat("\nmean:",mean(k[,1]))
k<-crossvalidate(data_train,Model="randomforest",methods="two-step")
cat("Model: randomforest\t\tMethods: two-step\n\n");k;cat("\nmean:",mean(k[,1]))
k<-crossvalidate(data_train,Model="ridge",methods="directly")
cat("Model: ridge regression\t\tMethods: directly\n\n");k;cat("\nmean:",mean(k[,1]))
k<-crossvalidate(data_train,Model="ridge",methods="two-step")
cat("Model: ridge regression\t\tMethods: two-step\n\n");k;cat("\nmean:",mean(k[,1]))
k<-crossvalidate(data_train,Model="lasso",methods="directly")
cat("Model: lasso\t\tMethods: directly\n\n");k;cat("\nmean:",mean(k[,1]))
k<-crossvalidate(data_train,Model="lasso",methods="two-step")
cat("Model: lasso\t\tMethods: two-step\n\n");k;cat("\nmean:",mean(k[,1]))

```

## Test Set Results
### Direct Prediction
```{r eval=T}
ranking_names<-c(data_test$name[data_test$rank_in_finals==1],
  data_test$name[data_test$rank_in_finals==2],
  data_test$name[data_test$rank_in_finals==3])
#########################################
########### Direct Prediction ###########
#########################################
```
#### Ridge Regression
```{r eval=T}
## Ridge Regression ##
r_ini<-glmnet(x=as.matrix(data_train[,-(1:4)]),y=data_train$rank_in_finals,alpha=0)
plotCoeffEvolution(r_ini)
ridge_cv<-cv.glmnet(x=as.matrix(data_train[,-(1:4)]),y=data_train$rank_in_finals,
                    type.measure="mse",foldid=data_train$seasons,alpha=0)
plot(ridge_cv)
cat("Lambda chosen by Mean-Square Error: ",ridge_cv$lambda.min)
score_ridge<-predict(ridge_cv, new=as.matrix(data_test[,-(1:4)]))
# cat("Scores:\n ",score_ridge)
k<-t(as.data.frame(score_ridge))
colnames(k)<-data_test$name
rownames(k)<-"Scores"
kable(k)
# score_ridge[rank(score_ridge)>3]<-10
rp<-rank(score_ridge,ties.method="first")
rts<-rank(data_test$rank_in_finals,ties.method="first")
corders<-sum(rts[which(rp<=3)]==rp[which(rp<=3)])
rank<-rbind(c(1,2,3),c(rp[data_test$rank_in_finals==1],
                       rp[data_test$rank_in_finals==2],rp[data_test$rank_in_finals==3]))
row.names(rank)<-c("Actual Ranking","Prediction")
colnames(rank)<-ranking_names
kable(rank)
cat("Accuracy of Prediction:",corders,"/ 3") #3/3
```

#### Lasso
```{r eval=T}
## Lasso ##
l_ini<-glmnet(x=as.matrix(data_train[,-(1:4)]),y=data_train$rank_in_finals,alpha=1)
plotCoeffEvolution(l_ini)
lasso_cv<-cv.glmnet(x=as.matrix(data_train[,-(1:4)]),y=data_train$rank_in_finals,
                    type.measure="mse",foldid=data_train$seasons,alpha=1)
plot(lasso_cv)
cat("Lambda chosen by Mean-Square Error: ",lasso_cv$lambda.min)
score_lasso<-predict(lasso_cv, new=as.matrix(data_test[,-(1:4)]))
# cat("Scores:\n ",score_lasso)
k<-t(as.data.frame(score_lasso))
colnames(k)<-data_test$name
rownames(k)<-"Scores"
kable(k)
rp<-rank(score_lasso,ties.method="first")
rts<-rank(data_test$rank_in_finals,ties.method="first")
corders<-sum(rts[which(rp<=3)]==rp[which(rp<=3)])
rank<-rbind(c(1,2,3),c(rp[data_test$rank_in_finals==1],
                       rp[data_test$rank_in_finals==2],rp[data_test$rank_in_finals==3]))
row.names(rank)<-c("Actual Ranking","Prediction")
colnames(rank)<-ranking_names
kable(rank)
cat("Accuracy of Prediction:",corders,"/ 3") #3/3
imp_lasso<-c(names(coef(lasso_cv)[,1])[-1][rank(-varImp(lasso_cv$glmnet.fit,lasso_cv$lambda.1se)[,1])==1],
             names(coef(lasso_cv)[,1])[-1][rank(-varImp(lasso_cv$glmnet.fit,lasso_cv$lambda.1se)[,1])==2],
             names(coef(lasso_cv)[,1])[-1][rank(-varImp(lasso_cv$glmnet.fit,lasso_cv$lambda.1se)[,1])==3])

```

#### Random Forest
```{r eval=T}
## Random Forest ##
rf_rank<-randomForest(x=data_train[,-(1:4)],y=data_train$rank_in_finals,ntree=500)
score_rf<-predict(rf_rank,newdata=data_test[,-(1:4)])
# cat("Scores:\n ",score_rf)
k<-t(as.data.frame(score_rf))
colnames(k)<-data_test$name
rownames(k)<-"Scores"
kable(k)
# score_rf[rank(score_rf)>3]<-10
rp<-rank(score_rf,ties.method="first")
rts<-rank(data_test$rank_in_finals,ties.method="first")
corders<-sum(rts[which(rp<=3)]==rp[which(rp<=3)])
rank<-rbind(c(1,2,3),c(rp[data_test$rank_in_finals==1],
                       rp[data_test$rank_in_finals==2],rp[data_test$rank_in_finals==3]))
row.names(rank)<-c("Actual Ranking","Prediction")
colnames(rank)<-ranking_names
kable(rank)
cat("Accuracy of Prediction:",corders,"/ 3")
# explain_forest(rf_rank)
imp_rf<-names(sort(rank(-rf_rank$importance[,1])))[1:3]
```

### Two-step Prediction
```{r eval=T}
#########################################
########## Two-step Prediction ##########
#########################################
```

#### Ridge Regression
```{r eval=T}
## Ridge Regression ##
model<-cv.glmnet(x=as.matrix(data_train[,-(1:4)]),y=data_train$top3_or_not,
                 type.measure="mse",nfolds=4,alpha=0)
trained<-data_train$top3_or_not
predicted<-predict(model,new=as.matrix(data_test[,-(1:4)]))
predicted<-as.numeric(rank(-predicted)<=3)
cat("Accuracy of Prediction on top3_or_not: ",sum(predicted[data_test$top3_or_not==1]),"/ 3")
model<-cv.glmnet(x=as.matrix(cbind(data_train[,-(1:4)],new=trained)),
                 y=data_train$rank_in_finals,
                 type.measure="mse",nfolds=4,alpha=0)
r_ini<-glmnet(x=as.matrix(cbind(data_train[,-(1:4)],new=trained)),
                  y=data_train$rank_in_finals,alpha=0)
plotCoeffEvolution(r_ini)
plot(model)
cat("Lambda chosen by Mean-Square Error: ",model$lambda.min)
predicted<-predict(model,new=as.matrix(cbind(data_test[,-(1:4)],new=predicted)))
# cat("Scores:\n ",predicted)
k<-t(as.data.frame(predicted))
colnames(k)<-data_test$name
rownames(k)<-"Scores"
kable(k)
# predicted[rank(predicted)>3]<-10
# MSPE<-mspe(data_test$rank_in_finals,predicted)
# RMSPE<-rmspe(data_test$rank_in_finals,predicted)
# MAPE<-mape(data_test$rank_in_finals,predicted)
rp<-rank(predicted,ties.method="first")
rts<-rank(data_test$rank_in_finals,ties.method="first")
corders<-sum(rts[which(rp<=3)]==rp[which(rp<=3)])
rank<-rbind(c(1,2,3),c(rp[data_test$rank_in_finals==1],
                       rp[data_test$rank_in_finals==2],rp[data_test$rank_in_finals==3]))
row.names(rank)<-c("Actual Ranking","Prediction")
colnames(rank)<-ranking_names
kable(rank)
cat("Accuracy of Prediction:",corders,"/ 3")
```

#### Lasso
```{r eval=T}
## Lasso ##
model<-cv.glmnet(x=as.matrix(data_train[,-(1:4)]),y=data_train$top3_or_not,
                 type.measure="mse",nfolds=4,alpha=1)
trained<-data_train$top3_or_not
predicted<-predict(model,new=as.matrix(data_test[,-(1:4)]))
predicted<-as.numeric(rank(-predicted)<=3)
cat("Accuracy of Prediction on top3_or_not: ",sum(predicted[data_test$top3_or_not==1]),"/ 3")
model<-cv.glmnet(x=as.matrix(cbind(data_train[,-(1:4)],new=trained)),
                 y=data_train$rank_in_finals,
                 type.measure="mse",nfolds=4,alpha=1)
l_ini<-glmnet(x=as.matrix(cbind(data_train[,-(1:4)],new=trained)),
              y=data_train$rank_in_finals,alpha=1)
plotCoeffEvolution(l_ini)
plot(model)
cat("Lambda chosen by Mean-Square Error: ",model$lambda.min)
predicted<-predict(model,new=as.matrix(cbind(data_test[,-(1:4)],new=predicted)))
# predicted[rank(predicted)>3]<-10
# cat("Scores:\n ",predicted)
k<-t(as.data.frame(predicted))
colnames(k)<-data_test$name
rownames(k)<-"Scores"
kable(k)
# MSPE<-mspe(data_test$rank_in_finals,predicted)
# RMSPE<-rmspe(data_test$rank_in_finals,predicted)
# MAPE<-mape(data_test$rank_in_finals,predicted)
rp<-rank(predicted,ties.method="first")
rts<-rank(data_test$rank_in_finals,ties.method="first")
corders<-sum(rts[which(rp<=3)]==rp[which(rp<=3)])
rank<-rbind(c(1,2,3),c(rp[data_test$rank_in_finals==1],
                       rp[data_test$rank_in_finals==2],rp[data_test$rank_in_finals==3]))
row.names(rank)<-c("Actual Ranking","Prediction")
colnames(rank)<-ranking_names
kable(rank)
cat("Accuracy of Prediction:",corders,"/ 3")
imp_lasso2<-c(names(coef(model)[,1])[-1][rank(-varImp(model$glmnet.fit,lasso_cv$lambda.min)[,1])==1],
             names(coef(model)[,1])[-1][rank(-varImp(model$glmnet.fit,lasso_cv$lambda.min)[,1])==2],
             names(coef(model)[,1])[-1][rank(-varImp(model$glmnet.fit,lasso_cv$lambda.min)[,1])==3])

```

#### Random Forest
```{r eval=T}
## Random Forest ##
model<-randomForest(data_train[,-(1:4)],data_train$top3_or_not,ntree=500)
trained<-predict(model)
trained<-data_train$top3_or_not
predicted<-predict(model,newdata=data_test[,-(1:4)])
predicted<-as.numeric(rank(-predicted)<=3)
cat("Accuracy of Prediction on top3_or_not: ",sum(predicted[data_test$top3_or_not==1]),"/ 3")
model<-randomForest(cbind(data_train[,-(1:4)],new=trained),
                    data_train$rank_in_finals,ntree=500)
predicted<-predict(model,newdata=cbind(data_test[,-(1:4)],new=predicted))
# predicted[rank(predicted)>3]<-10
# cat("Scores:\n ",predicted)
k<-t(as.data.frame(predicted))
colnames(k)<-data_test$name
rownames(k)<-"Scores"
kable(k)
# MSPE<-mspe(data_test$rank_in_finals,predicted)
# RMSPE<-rmspe(data_test$rank_in_finals,predicted)
# MAPE<-mape(data_test$rank_in_finals,predicted)
rp<-rank(predicted,ties.method="first")
rts<-rank(data_test$rank_in_finals,ties.method="first")
corders<-sum(rts[which(rp<=3)]==rp[which(rp<=3)])
rank<-rbind(c(1,2,3),c(rp[data_test$rank_in_finals==1],
                       rp[data_test$rank_in_finals==2],rp[data_test$rank_in_finals==3]))
row.names(rank)<-c("Actual Ranking","Prediction")
colnames(rank)<-ranking_names
kable(rank)
cat("Accuracy of Prediction:",corders,"/ 3")
imp_rf2<-names(sort(rank(-model$importance[,1])))[1:3]
```

```{r eval=T}

imp_tb<-data.frame(imp_lasso,imp_rf,imp_lasso2,imp_rf2)
colnames(imp_tb)<-c("lasso directly","rforest directly","lasso two-step","rforest two-step")
cat("Features Importance Given By Lasso and Random Forest:\n");kable(imp_tb)

```

